---
title: "Forecast of COVID-19 Cases in LA Cities Based on Cluster"
output: html_document
runtime: shiny
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r Libraries, include = FALSE}
library(forecast)
library(ggplot2)
library(gridExtra)
library(tidyverse)
```


```{r Cluster Cases, include = FALSE}
LA_cluster_cases = read.csv(file = "LA_cities_cases.csv", header = FALSE)
LA_cluster = read.csv(file = "new_LA_cities_clusters.csv", header = FALSE)

case_cluster = merge(LA_cluster_cases, LA_cluster, by="V1", all=TRUE)
case_cluster = case_cluster %>% rename("Cluster" = "V2.y")
case_cluster = case_cluster[-c(1, 2), ]
case_cluster$V1 = NULL
case_cluster = case_cluster %>% mutate_if(is.character,as.numeric)

#sum of all cases in a cluster was used
case_cluster = aggregate(. ~ Cluster, data=case_cluster, FUN=sum)

df = t(case_cluster)
colnames(df) <- as.character(unlist(df[1,]))
df = df[-1, ]
df = as.data.frame(df)
```



```{r Severity Data, include = FALSE}
new_LA = read.csv(file = "new_LA_cities_severity.csv", header = FALSE)
LA_cluster = read.csv(file = "new_LA_cities_clusters.csv", header = FALSE)
sev_cluster = merge(new_LA, LA_cluster, by="V1", all=TRUE)
sev_cluster = sev_cluster %>% rename("Cluster" = "V2.y")
sev_cluster = sev_cluster[-c(1, 2), ]
sev_cluster$V1 = NULL
sev_cluster = sev_cluster %>% mutate_if(is.character,as.numeric)

#severity value mean was used for each cluster
sev_cluster = aggregate(. ~ Cluster, data=sev_cluster, FUN=mean)

severity_df = sev_cluster
severity_df = t(severity_df)
colnames(severity_df) <- as.character(unlist(severity_df[1,]))
severity_df = severity_df[-1, ]
severity_df = as.data.frame(severity_df)
```


```{r City Prediction Function, include = FALSE}
city_prediction_reg = function(city) {
  #Prediction with severity values as regressors
  city_timeseries = as.numeric(ts(df[,city]))
  
  #Fit to arima
  city_Severity = as.numeric(ts(severity_df[,city]))
  city_fit = try(auto.arima(city_timeseries, xreg = city_Severity), silent = TRUE)
  city_Sev_Pred = as.numeric(forecast(city_Severity)$mean) #prediction for future city_Severity values
  
  
  if (class(city_fit) == "try-error") {
    
    return(NULL)
    
  } else {
    
    #forecast with regressors
    city_Sev_Pred = as.numeric(forecast(city_Severity)$mean) #prediction for future city_Severity values
    
    return(autoplot(forecast(city_fit, xreg = city_Sev_Pred)) +
             xlab("Days Since Jan 21st") +
             ylab("Covid Cases")
    )
    
  }

}

city_prediction_nreg = function(city) {
  #Prediction without severity values as regressors
  city_timeseries = as.numeric(ts(df[,city]))
  city_fit = auto.arima(city_timeseries)

  #forecast no regressors  
  return(autoplot(forecast(city_fit)) +
           xlab("Days Since Jan 21st") +
           ylab("Covid Cases")
  )
}

city_prediction = function(city) {
  #complete city prediction function
  
  tryCatch(graph_compare,
    error = function(c) {
      #Sometimes an arima model can't be found with the severity values, so an arima model w/o severity regressors is used instead
      city_prediction_nreg(city)
      }
    
    )
  
}
```


```{r Error Calculations Table, include = FALSE}
city_pred_reg_error = function(city) {
  #prediction and error checking using severity values
  #taking first few data points and using it to train model
  first_95 = head(df[,city], length(df[,city]) - 10)
  first_95ts = ts(as.numeric(first_95))
  sev_first_95 = head(severity_df[,city], length(severity_df[,city]) - 10)
  sev_first_95ts = ts(as.numeric(sev_first_95))

  #Model trained on first days
  fit95 = auto.arima(first_95ts, xreg = sev_first_95ts)
  sev_95pred = as.numeric(forecast(sev_first_95ts)$mean)

  #predict last 10 days of dataset
  pred10 = as.data.frame(as.numeric(forecast(fit95, xreg = sev_95pred)$mean))   
  #actual values for last 10 days
  last_10_city = as.data.frame(as.numeric(tail(df[,city], 10)))

  #error
  errors = (pred10 - last_10_city)/last_10_city
  colnames(errors) = "Errors w/ Severity Values"
  return(errors)
}

city_pred_nreg_error = function(city) {
  #prediction and error checking without severity values
  #taking first few data points and using it to train model
  first_95 = head(df[,city], length(df[,city]) - 10)
  first_95ts = ts(as.numeric(first_95))
  
  #Model trained on first days
  fit95 = auto.arima(first_95ts)

  #predict last 10 days of dataset
  pred10 = as.data.frame(as.numeric(forecast(fit95)$mean))

  #actual values for last 10 days
  last_10_city = as.data.frame(as.numeric(tail(df[,city], 10)))

  #error
  errors = (pred10 - last_10_city)/last_10_city
  colnames(errors) = "Errors w/o Severity Values"
  return(errors)
}

city_pred_error = function(city) {
  
  sev_reg = try(city_pred_reg_error(city), silent = TRUE)
  n_sev_reg = city_pred_nreg_error(city)
  pred_days = as.data.frame(c(1:10))
  colnames(pred_days) = "Days Since End of Given Data - Prediction Days"

  errors_table = tryCatch(
    #if sev_reg not found because no suitable arima model then only finds errors with no severity values
    cbind(pred_days, sev_reg, n_sev_reg),
    
    error = function(err) {
      cbind(pred_days, n_sev_reg)
    }
    
  )
  
  return(errors_table)

}
```

```{r Cumulative City Prediction Function, include = FALSE}
cum_city_prediction_reg = function(city) {
  #Prediction with severity values as regressors
  city_timeseries = as.numeric(ts(cumsum(df[,city])))
  
  #Fit to arima
  city_Severity = as.numeric(ts(severity_df[,city]))
  city_fit = auto.arima(city_timeseries, xreg = city_Severity)
  
  #forecast with regressors
  city_Sev_Pred = as.numeric(forecast(city_Severity)$mean) #prediction for future city_Severity values
  
  autoplot(forecast(city_fit, xreg = city_Sev_Pred)) +
    xlab("Days Since Jan 21st") +
    ylab("Covid Cases")
}

cum_city_prediction_nreg = function(city) {
  #Prediction without severity values as regressors
  city_timeseries = as.numeric(ts(cumsum(df[,city])))
  city_fit = auto.arima(city_timeseries)

  #forecast no regressors  
  autoplot(forecast(city_fit)) +
    xlab("Days Since Jan 21st") +
    ylab("Covid Cases")
}

city_prediction = function(city) {
  #complete city prediction function
  
  tryCatch(city_prediction_reg(city),
    error = function(c) {
      #Sometimes an arima model can't be found with the severity values, so an arima model w/o severity regressors is used instead
      city_prediction_nreg(city)
      }
    
    )
  
}
```

```{r Error Calculations Table Cumulative, include = FALSE}
cum_city_pred_reg_error = function(city) {
  #prediction and error checking using severity values
  #taking first few data points and using it to train model
  first_95 = head(cumsum(df[,city]), length(df[,city]) - 10)
  first_95ts = ts(as.numeric(first_95))
  sev_first_95 = head(severity_df[,city], length(severity_df[,city]) - 10)
  sev_first_95ts = ts(as.numeric(sev_first_95))

  #Model trained on first days
  fit95 = auto.arima(first_95ts, xreg = sev_first_95ts)
  sev_95pred = as.numeric(forecast(sev_first_95ts)$mean)

  #predict last 10 days of dataset
  pred10 = as.data.frame(as.numeric(forecast(fit95, xreg = sev_95pred)$mean))   
  #actual values for last 10 days
  last_10_city = as.data.frame(as.numeric(tail(cumsum(df[,city]), 10)))

  #error
  errors = (pred10 - last_10_city)/last_10_city
  colnames(errors) = "Errors w/ Severity Values"
  return(errors)
}

cum_city_pred_nreg_error = function(city) {
  #prediction and error checking without severity values
  #taking first few data points and using it to train model
  first_95 = head(cumusum(df[,city]), length(df[,city]) - 10)
  first_95ts = ts(as.numeric(first_95))
  
  #Model trained on first days
  fit95 = auto.arima(first_95ts)

  #predict last 10 days of dataset
  pred10 = as.data.frame(as.numeric(forecast(fit95)$mean))

  #actual values for last 10 days
  last_10_city = as.data.frame(as.numeric(tail(cumsum(df[,city]), 10)))

  #error
  errors = (pred10 - last_10_city)/last_10_city
  colnames(errors) = "Errors w/o Severity Values"
  return(errors)
}

cum_city_pred_error = function(city) {
  
  sev_reg = try(city_pred_reg_error(city), silent = TRUE)
  n_sev_reg = city_pred_nreg_error(city)

  errors_table = tryCatch(
    #if sev_reg not found because no suitable arima model then only finds errors with no severity values
    cbind(sev_reg, n_sev_reg),
    
    error = function(err) {
      n_sev_reg
    }
    
  )
  
  return(errors_table)

}
```

```{r testing, include = FALSE}
graph_compare = function(city) {
  #trying to rearrange graphs side by side
  par(mfrow=c(1,2))
  city_prediction_reg(city)
  city_prediction_nreg(city)
}
  
test1 = city_prediction_reg("1")
test2 = city_prediction_nreg("1")
grid.arrange(test1, test2, nrow = 1,
             widths = c(2.7, 1))


```

## Covid Forecasts by Cluster

Hierarchical clustering was used to group cities into 16 clusters. Regression models with severity values and ARIMA errors can be found for most clusters. For some clusters however, a proper ARIMA model using severity values can't be found, so only a normal ARIMA model is given.


```{r Inputs, echo=FALSE}
inputPanel(
  selectInput(inputId = "city", 
                label = "Cluster:",
                choices = colnames(df)
                )
  )
```

```{r Plot Forecasts, echo = FALSE}
renderPlot(city_prediction_reg(input$city))
renderPlot(city_prediction_nreg(input$city))
```


## Prediction Errors of Last 10 Days

Each model predicts the final 10 days of the given data set. The predicted values are compared to the true values, and their errors are found with the formula $\frac{Predicted-Actual}{Actual}$.

```{r Errors Tables, echo = FALSE}
renderDataTable(city_pred_error(input$city),
                options = list(dom = 't'))
```

